---
title: "GuideToOptizyme"
author: "Michelle Awh"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GuideToOptizyme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Optizyme

 <http://github.com/genehackers/optizyme>

Optizyme is a software package in R that streamlines the optimization and modelling of cell-free biological systems.

Optizyme is designed to be accessible for biologists without a heavy background in mathematics, computational methods in biology, or computer skills.

This document will serve as a tutorial to using the capabilities of Optizyme. Below we will outline the capabilities of Optizyme and detail how to use each of the functions included in the package. 

The functions included are:

1. **MCGD**: Optimization of enzyme ratios
2. **EnzymeModel**: Construction of a model for use with MCGD
3. **EnzymeModelPlot**: Visualization of the time course for a cell-free system
4. **EnzymeModelTimeCourse**: Ability to retrieve time course data for a more flexible visualization

```{r setup}

```
